---
title: "The 'I' in Team"
subtitle: "Peer-to-peer best practices for growing data science teams"
author: 
  - name: Liz Roten
    orcid: 0000-0002-5346-3549
    email: liz.roten@metc.state.mn.us
    affiliations:
      - name: Metropolitan Council
        address: 390 Robert St. N.
        city: Saint Paul
        state: MN
        postal-code: 55101
date: "September 19, 2023"
date-format: long
format: 
  revealjs:
    header: "posit::conf(2023)"
    footer: '<a href="https://lizroten.com/team" target="_blank">lizroten.com/team</a>'
    include-after-body: partials/clean_title_page.html
    hide-from-titleSlide: "all"
    menu: false
    template: partials/template.html
    theme: 
      - theme.scss
      - partials/title-slide.scss
    incremental: true
    standalone: true
    mouse-wheel: true
    template-partials:
      - partials/title-slide.html
      - partials/toc-slide.html
filters: 
  - reveal-header
include-in-header:
  - text: |
      <link rel = "shortcut icon" href = "https://em-content.zobj.net/source/google/350/busts-in-silhouette_1f465.png" />
---
```{r}
# window palette
harem <- list(
  prussian_blue      = "#0B273C",
  cadet_gray         = "#989F9C",
  carribbean_current = "#01606E",
  pale_dogwood       = "#E3C0B6",
  cadet_gray2        = "#A1B4B9",
  bistre             = "#2E180F",
  chinese_violet     = "#856084",
  gold               = "#A68247"
)


window_pal <- list(
  dark_blue =   "#162059",
  med_blue =    "#383E45",
  light_blue =  "#727A8C",
  yellow =      "#F2D091",
  tan =         "#D9CBBF",
  red =         "#592816",
  dark_red =    "#33190F")

library(ggplot2)
library(magrittr)
library(councilR)

showtext::showtext_auto()
sysfonts::font_paths()
files <- sysfonts::font_files()
sysfonts::font_add("HelveticaNeueLT Std Cn", "HelveticaNeueLTStd-Cn.otf")
sysfonts::font_add("HelveticaNeueLT Std Lt", "HelveticaNeueLTStd-Lt.otf")
sysfonts::font_add("HelveticaNeueLT Std Thin", "HelveticaNeueLTStd-Th.otf")
sysfonts::font_add("Arial Narrow", "Arial Narrow.ttf")
sysfonts::font_add("HelveticaNeueLT Std Med Cn", "HelveticaNeueLTStd-MdCn.otf")
sysfonts::font_add("Palatino Linotype", "pala.ttf")
```


## 

In my workplace, we started off with a tiny data science team. 

<!-- need *time*, place, and characters -->
. . .


And by tiny I mean 2 people. 

Over time, we had some turnover and our team grew. 

. . . 

we managed to expand our team, and teach more folks about R.

. . . 

when we got more data scientists on board, 

. . .

it became clear that we were all **_very_** used to working alone with a single incident. 

. . . 

## ...how bad was this incident?

I maintain an R package for folks in my workplace.

<!-- It has all sorts of practical things, like color codes, ggplot2 themes, and database helpers.  -->

Of my own accord, I decided that we should use a more flexible font sizing scheme for our ggplot2 theme. 

I made the changes, released the new version, and encouraged my colleagues to install it. 

All was going well, until my colleague tried to actually use the new functions.

...

... while they were writing a time sensitive report

## 

```{r fig-good-plot}
#| fig-cap-location: top
#| fig-height: 6
color_range_blue <- colorRampPalette(c("#e0e3e8", colors$playDaBlue))

ggplot(datasets::airquality) +
  aes(Solar.R, Wind, 
      color = as.factor(Month)) +
  geom_point(size = 9) +
  scale_color_manual(values = color_range_blue(5)) +
  labs(title = "A very important plot",
       color = "Month",
       x = "Solar",
       y = "Wind",
       caption = stringr::str_wrap(shinipsum::random_text(nwords = 20))) +
  theme_council(use_showtext = TRUE,
                base_size = 30) +
  theme(plot.caption = element_text(size = 10),
        plot.caption.position = "plot")
```

## 

```{r fig-bad-plot}
#| fig-cap-location: top
#| fig-height: 6

multiplier <- 1.5

ggplot(datasets::airquality) +
  aes(Solar.R, Wind, 
      color = as.factor(Month)) +
  geom_point(size = 9) +
  scale_color_manual(values = color_range_blue(5)) +
  labs(title = "A very important plot",
       color = "Month",
       x = "Solar",
       y = "Wind",
       caption = stringr::str_wrap(shinipsum::random_text(nwords = 20))) +
  theme_council(use_showtext = TRUE,
                use_manual_font_sizes = TRUE,
                font_sizes = list(
                            "title" = 20 * multiplier,
                            "subtitle" = 16 * multiplier,
                            "axis_title" = 16 * multiplier,
                            "axis_text" = 6 * multiplier,
                            "legend_title" = 9 * multiplier,
                            "legend_text" = 14 * multiplier,
                            "caption" = 16 * multiplier,
                            "strip" = 14 * multiplier
                          )) +
  theme(axis.title.y = element_text(angle = 90),
        plot.title = element_text(family = "HelveticaNeueLT Std Med Cn"),
        plot.caption = element_text(
          family = "Palatino Linotype"
        ))

```
## So what went wrong?

- Innacurate and incomplete documentation
- Installed without reading docs
- 

##
<!-- This isn't exactly ideal for building a functional team -->
<!-- I don't think this story is particularly unique -->
<!-- we all make mistakes that break other peoples' code -->
- with shared projects
- 



## 



this talk is not about being a manager or how to manage a data science team, 

but how to be a good colleague when someone new comes in

## 

<!-- Work is work, and what makes it enjoyable, more than the subject matter, is the *people* you get work with and the team you create.  -->

Your manager will accomplish lots of important things for bringing in a new person - but you are the person they get to work with day-to-day!



## 

1. Make the implicit explicit
1. All knowledge is valid and valuable
1. Document (aka, make life easier for your friends)

# Make the implicit explicit {background-color="#A1B4B9"}



## Workplace culture {background-color="#A1B4B9"}


## {background-color="#A1B4B9"}

don't
- Everyone thinks that they don't have culture, but really, its what we are all surrounded with constantly.
- Start by making a list of the tools you use on a daily basis, no matter how big or small
<!-- go through how you use them, and to what extent -->
<!-- how do they make your life easier? -->
- I recommend using a Git GUI (like GitKraken)
<!-- - A git GUI is a great way to visualize branches and track progress over time. -->
<!-- It also  -->




# All knowledge is valuable {background-color="#D9CBBF"}

## {background-color="#D9CBBF"}


they have certain skill your team may have been lacking

Data science is a deeply creative field. You can hand two different people a challenging problem and they both can come up with totally different ways of solving it

all the knowledge that the new person is bringing in 

There is a lot of work that your manager will take care of...

the recipe for trust 


##  {background-color="#D9CBBF"}

- Patience will prove best
<!-- your manager will do a lot of the legwork to get this person onboard,  -->
<!-- like setting up timesheets and going through workplace policies -->
<!-- but you are still the person they will be actually working with -->
<!-- you will have to bring this person up to speed, and that is part of it -->
- You(r manager) hired this person for a reason
-
# Document {background-color="#E3C0B6"}


## A matter of investment {background-color="#E3C0B6"}


- **Invest** in tools that encourage documentation and lower cognitive load.
- **Invest** in the time it takes to document/clean up/organize your work
- **Invest** in strong social relationships

<!-- do things that don't feel like work -->
<!-- - Try making every other team meeting a casual hang-out -->
- Schedule in time to build connections and trust. These social gatherings are just as valuable as focus time


<!-- Even if your new colleague says they have experience with git, that could mean very different things.  -->


# Ultimately...

## 

I've grown immensely from working with my colleagues

Investing in your team not just as a manger, but as a coworker

# Thank you

<a href="https://lizroten.com/gracefully" target="_blank">lizroten.com/gracefully</a>
